---
title: "Data Reading and Preprocessing"
author: "Nashra Ahmad"
date: '2024-01-28'
output: 
  html_document:
    toc: true
    number_sections: true
    theme: flatly
    highlight: espresso
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up libraries

```{r libraries,warning=FALSE,message=FALSE}
library(tidyverse)
library(ggplot2)
library(stringr)
#library(sjstats) # TE: This library is not used 
```

## Read data

Read Western data

```{r readdata}
d1<-read.csv('DurhamMusicFinal.csv',skip = 0,header = TRUE)
d2<-read.csv('DurhamBs1.csv',skip = 0,header = TRUE)
d3<-read.csv('DurhamBsFinal.csv',skip = 0,header = TRUE)
d4<-read.csv('DurhamUni.csv',skip = 0,header = TRUE)
d5<-read.csv('RhythmLearning_Study1 - Prolific_December 30, 2023_12.09.csv',skip = 0,header = TRUE)

d1<-d1[3:nrow(d1),]# delete two first rows (metadata)
d2<-d2[3:nrow(d1),]# TE: add this well
d3<-d3[3:nrow(d1),]
d4<-d4[3:nrow(d1),]
d5<-d5[3:nrow(d1),]
dw <- rbind(d1, d2, d3, d4, d5)
#df<-d[3:nrow(d),]# TE: removed
print(dim(dw))
```

Read Indian data

```{r readdataIndian}
d6<-read.csv('For Indians_RhythmLearning_Study1_January 28, 2024_12.21.csv',skip = 0,header = TRUE)
d6<-d6[3:nrow(d6),]# delete two first rows (metadata)
print(dim(d6))
```


## Filter data

Western data

```{r kill-rows}
dw <- dplyr::filter(dw,as.numeric(Progress) > 96)
dw <- dplyr::filter(dw,as.numeric(Duration..in.seconds.)>4*60) # 4 mins
print(nrow(dw))
```

Indian data

```{r kill-rows2}
di<- dplyr::filter(d6,as.numeric(Progress) > 96)
di <- dplyr::filter(di,as.numeric(Duration..in.seconds.)>4*60) # 4 mins
print(nrow(di))
```

## Reporting age and sd of the dataset

western data

```{r age}
numeric_age <- as.numeric(dw$Q_Age)
mean_age <- mean(numeric_age, na.rm = TRUE) #age of participants
sd_age <- sd(numeric_age, na.rm = TRUE) #sd for age
print(paste('Unfamiliar (Western) Mean age:',round(mean_age,2), 'SD:',round(sd_age,2)))
```

Indian data

```{r age2}
numeric_age2 <- as.numeric(di$Q_Age)
mean_age2 <- mean(numeric_age2, na.rm = TRUE) #age of participants
sd_age2 <- sd(numeric_age2, na.rm = TRUE) #sd for age
print(paste('Familiar (Indian) Mean age:',round(mean_age2,2), 'SD:',round(sd_age2,2)))
```

## Removing Indian participants from Non-Indian data

```{r group1,results='asis'}
dw$Familiarity <- ifelse(dw$Q_Ethnicity %in% c('Indian', 'Asian - Indian', 'Asian -Indian','Asian- Indian','Asian-Indian','Indian ', 'Asian - Indian '), 'Indian Ethnicity', 'Non-Indian Ethnicity')
Sf<-summarise(group_by(dw,Familiarity),n=n()) # count
knitr::kable(Sf)
```

```{r}
#(optional)writing Indian Participants seperately
#subsetind_df <- df[df$Familiarity == 'Indian Ethnicity', ]
# Write the subset to a CSV file
#write.csv(subsetind_df, file = "indian_ethnicity.csv", row.names = FALSE)
```

##Group into musicians and non-musicians Western data

```{r groupwes,results='asis'}
dw$Musicianship <- ifelse(dw$Q_Musicianship %in% c('Amateur musician', 'Serious amateur musician','Semiprofessional musician', 'Professional musician'), 'Musician', 'Non-Musician')
dw<-drop_na(dw) # drop missing observations 
Sm<-summarise(group_by(dw,Musicianship,Familiarity),n=n()) # count
knitr::kable(Sm)
```

```{r group3,results='asis'}
#deleting Indian participants from western data
dw1 <- dw[!(dw$Musicianship == 'Musician' & dw$Familiarity == 'Indian Ethnicity') & !(dw$Musicianship == 'Non-Musician' & dw$Familiarity == 'Indian Ethnicity') , ]
Sfilter<-summarise(group_by(dw1,Musicianship,Familiarity),n=n()) # count
knitr::kable(Sfilter)
```

Indian data

```{r musicianship in}
# Create a new column 'Musicianship' based on Score
di$Musicianship <- ifelse(di$ScoreFam >= 4, 'Musician', ifelse(is.na(di$ScoreFam) | di$ScoreFam < 4, 'Non-Musician', NA))
Sm2<-summarise(group_by(di,Musicianship),n=n()) # count
knitr::kable(Sm2)
```

##Arranging Columns

```{r kill-columns3}
#western data
dw2 <- dplyr::select(dw1,ResponseId,Musicianship,Familiarity, starts_with('Base_'), starts_with('Test_'))
dw_long <- tidyr::pivot_longer(dw2,ends_with('_1'))#changing to long

#indian data
di2 <- dplyr::select(di,ResponseId,Musicianship,starts_with('Base_'), starts_with('Test_'))
di_long <- tidyr::pivot_longer(di2,ends_with('_1'))#changing to long
```

## Code the factor names and rating values

Western data

```{r recode1}
dw_long$condition <- str_split_i(dw_long$name,'_',1)#splittig condition Baseline/Testing
#print (df_long$condition)
dw_long$pattern <- str_split_i(dw_long$name,'_',2)#splittig pattern Teental/Rupak
#print (df_long$pattern)
dw_long$type <- str_split_i(dw_long$name,'_',3) #splitting type of rhythm sd,sp,natural,basic,complex
#print (df_long$type)   

table(dw_long$condition)
table(dw_long$pattern)
table(dw_long$type)
table(dw_long$type,dw_long$condition)
```

Indian data

```{r recode2}
di_long$condition <- str_split_i(di_long$name,'_',1)#splittig condition Baseline/Testing
#print (df_long$condition)
di_long$pattern <- str_split_i(di_long$name,'_',2)#splittig pattern Teental/Rupak
#print (df_long$pattern)
di_long$type <- str_split_i(di_long$name,'_',3) #splitting type of rhythm sd,sp,natural,basic,complex
#print (df_long$type)   

# TE adds these reports just to check what is in the data (from previous code)
table(di_long$condition)
table(di_long$pattern)
table(di_long$type)
table(di_long$type,di_long$condition) 
```

#Converting Strings to Numbers Western data

```{r recode3}
table(dw_long$value)# since the ratings are string, not numbers, convert them
dw_long$value_c <- factor(dw_long$value,levels = rev(c("Extremely Similar","Very Similar","Similar","Moderately Similar","Slightly Similar","Not Similar","Not at all similar")))
table(dw_long$value_c)


dw_long$similarity <- as.numeric(dw_long$value_c) # as numeric
table(dw_long$similarity)
```

Indian data

```{r recode4}
table(di_long$value)# since the ratings are string, not numbers, convert them
di_long$value_c <- factor(di_long$value,levels = rev(c("Extremely Similar","Very Similar","Similar","Moderately Similar","Slightly Similar","Not Similar","Not at all similar")))
table(di_long$value_c)
di_long$similarity <- as.numeric(di_long$value_c) # as numeric
table(di_long$similarity)
```

#Checking the number of observations Western data

```{r check1,results='asis',warning=FALSE,message=FALSE}
dw_long<-drop_na(dw_long) # drop missing observations
S<-summarise(group_by(dw_long,pattern,condition),n=n()) # count
knitr::kable(S)
```

Indian data

```{r check2,results='asis',warning=FALSE,message=FALSE}
di_long<-drop_na(di_long) # drop missing observations
S2<-summarise(group_by(di_long,pattern,condition),n=n()) # count
knitr::kable(S2)
```

##writing both raw data seperately for later analysis (if needed)

```{r rewrite1}
# Write the subset to a CSV file
write.csv(dw_long, file = "Westerndata.csv", row.names = FALSE)
write.csv(di_long, file = "Indiandata.csv", row.names = FALSE)
```

#quality control: Western data

##quality control question

```{r quality1}
#Explanation of naming in experiment: Naturl= Natural for both Teental and Rupak, Naturll= Test for Teental, Naturl2= Test for Rupak

# Subset for 'Teental' and 'Rupak'
dw_subset_teental <- dw_long[(dw_long$pattern == 'Teental'& dw_long$condition=='Test' & (dw_long$type == 'Naturl' | dw_long$type == 'Naturll')),]
dw_subset_rupak <- dw_long[(dw_long$pattern == 'Rupak'& dw_long$condition=='Test' & (dw_long$type == 'Naturl' | dw_long$type == 'Naturl2')), ]

dw_subteental <- dw_subset_teental %>%
  filter(type %in% c('Naturll', 'Naturl')) %>%
  group_by(ResponseId) %>%
  summarise(AbsDifference = abs(diff(similarity[type %in% c('Naturl', 'Naturll')])))

# Direction of the difference is meaningless, so use absolute instead

dw_subrupak <- dw_subset_rupak %>%
  filter(type %in% c('Naturl', 'Naturl2')) %>%
  group_by(ResponseId) %>%
  summarise(AbsDifference = abs(diff(similarity[type %in% c('Naturl', 'Naturl2')])))

#plotting teental
hist(dw_subteental$AbsDifference)
#plotting rupak
hist(dw_subrupak$AbsDifference)

dw_subrupak$ResponseId[dw_subrupak$AbsDifference >2]# also display the id of the person who is inconsistent
dw_subteental$ResponseId[dw_subteental$AbsDifference >2]# also display the id of the person 
```

## Cronbach's alpha

western musicians

```{r wesalpha}
library(psych)

dwmus <- dw_long[dw_long$Musicianship=='Musician' & dw_long$type %in% c('sd1', 'sp1', 'Basic', 'Naturl', 'Complex'),]
U <- unique(dwmus$type)
S <- unique(dwmus$ResponseId)
alpha <- matrix(0, 1, length(U))

TMP <- NULL
for (k in 1:length(S)) {
  B <- dplyr::filter(dwmus, ResponseId == S[k])
  TMP <- cbind(TMP, as.numeric(B$similarity))
}
colnames(TMP) <- S
a1 <- suppressMessages(suppressWarnings(psych::alpha(TMP, check.keys = FALSE, warnings = FALSE)))
Alpha1 <- a1$total$raw_alpha

# display individual ones
print(a1$item.stats$r.cor)

colnames(TMP)[a1$item.stats$r.cor < 0] # anybody under 0?
colnames(TMP)[a1$item.stats$r.cor < 0.7]
# Calculate mean alpha
mean_alpha1 <- mean(Alpha1)
print(mean_alpha1)

```


Western non-musicians

```{r alpha continued2}
dwmus <- dw_long[dw_long$Musicianship=='Non-Musician' & dw_long$type %in% c('sd1', 'sp1', 'Basic', 'Naturl', 'Complex'),]
U <- unique(dwmus$type)
S <- unique(dwmus$ResponseId)
alpha <- matrix(0, 1, length(U))

TMP <- NULL
for (k in 1:length(S)) {
  B <- dplyr::filter(dwmus, ResponseId == S[k])
  TMP <- cbind(TMP, as.numeric(B$similarity))
}
colnames(TMP) <- S
a2 <- suppressMessages(suppressWarnings(psych::alpha(TMP, check.keys = FALSE, warnings = FALSE)))
Alpha2 <- a2$total$raw_alpha

# display individual ones
a2$item.stats$r.cor

colnames(TMP)[a2$item.stats$r.cor < 0]
colnames(TMP)[a2$item.stats$r.cor < 0.7]
# Calculate mean alpha
mean_alpha2 <- mean(Alpha2)
print(mean_alpha2)
```

[NA: all three participants (R_2Uax5iSyFsc52pZ, R_3kpjZD1SAq9rBt8,
R_3lYS6mLft5j6zOW) participants with alpha \<0.7 match with the one in
quality control question test above, should we remove the
data?]{style="color: red"}

#quality control: Indian data ##quality control question

```{r 1}
#Explanation of naming in experiment: Naturl= Natural for both Teental and Rupak, Naturll= Test for Teental, Naturl2= Test for Rupak

# Subset for 'Teental' and 'Rupak'
di_subset_teental <- di_long[(di_long$pattern == 'Teental'& di_long$condition=='Test' & (di_long$type == 'Naturl' | di_long$type == 'Naturll')),]
di_subset_rupak <- di_long[(di_long$pattern == 'Rupak'& di_long$condition=='Test' & (di_long$type == 'Naturl' | di_long$type == 'Naturl2')), ]

di_subteental <- di_subset_teental %>%
  filter(type %in% c('Naturll', 'Naturl')) %>%
  group_by(ResponseId) %>%
  summarise(AbsDifference = abs(diff(similarity[type %in% c('Naturl', 'Naturll')])))

# Direction of the difference is meaningless, so use absolute instead

di_subrupak <- di_subset_rupak %>%
  filter(type %in% c('Naturl', 'Naturl2')) %>%
  group_by(ResponseId) %>%
  summarise(AbsDifference = abs(diff(similarity[type %in% c('Naturl', 'Naturl2')])))

#plotting teental
hist(di_subteental$AbsDifference)
#plotting rupak
hist(di_subrupak$AbsDifference)

di_subrupak$ResponseId[di_subrupak$AbsDifference >2]# also display the id of the person who is inconsistent
di_subteental$ResponseId[di_subteental$AbsDifference >2]# also display the id of the person 


```

## Cronbach's alpha

Indian Musicians

```{r 2}
# subsetting data for musicians
dimus <- di_long[di_long$Musicianship=='Musician' & di_long$type %in% c('sd1', 'sp1', 'Basic', 'Naturl', 'Complex'),]
U <- unique(dimus$type)
S <- unique(dimus$ResponseId)
alpha <- matrix(0, 1, length(U))

TMP <- NULL
for (k in 1:length(S)) {
  B <- dplyr::filter(dimus, ResponseId == S[k])
  TMP <- cbind(TMP, as.numeric(B$similarity))
}
colnames(TMP) <- S
a3 <- suppressMessages(suppressWarnings(psych::alpha(TMP, check.keys = FALSE, warnings = FALSE)))
Alpha3 <- a3$total$raw_alpha

# display individual ones
a3$item.stats$r.cor

colnames(TMP)[a3$item.stats$r.cor < 0] # anybody under 0?

colnames(TMP)[a3$item.stats$r.cor < 0.7]
# Calculate mean alpha
mean_alpha3 <- mean(Alpha3)
print(mean_alpha3)
```

Indian Non-Musician

```{r alpha continued3}
dimus <- di_long[di_long$Musicianship=='Non-Musician' & di_long$type %in% c('sd1', 'sp1', 'Basic', 'Naturl', 'Complex'),]
U <- unique(dimus$type)
S <- unique(dimus$ResponseId)
alpha <- matrix(0, 1, length(U))

TMP <- NULL
for (k in 1:length(S)) {
  B <- dplyr::filter(dimus, ResponseId == S[k])
  TMP <- cbind(TMP, as.numeric(B$similarity))
}
colnames(TMP) <- S
a4 <- suppressMessages(suppressWarnings(psych::alpha(TMP, check.keys = FALSE, warnings = FALSE)))
Alpha4 <- a4$total$raw_alpha

# display individual ones
a4$item.stats$r.cor

colnames(TMP)[a4$item.stats$r.cor < 0]
colnames(TMP)[a4$item.stats$r.cor < 0.7]
# Calculate mean alpha
mean_alpha4 <- mean(Alpha4)
print(mean_alpha4)
```

[NA: 4 of the participants (R_4PhKWMzINpHFoVy, R_4Y9utJ6hhxLFuDy,
R_4eWPin0vYILik8E, R_4qsOSprsTyIzn7e) with alpha \<0.7 match with the
one in quality control question test above, should we remove that
data?]{style="color: red"}

##Checking  Participants getting Removed
```{r checking}
# Western
dw2 <- dw[!(dw$ResponseId %in% c('R_2Uax5iSyFsc52pZ', 'R_3kpjZD1SAq9rBt8', 'R_3lYS6mLft5j6zOW')), ]
#dw_long <- dw_long[!(dw_long$ResponseId %in% c('R_2Uax5iSyFsc52pZ', 'R_3kpjZD1SAq9rBt8', 'R_3lYS6mLft5j6zOW')), ]
# Checking the loss
Sfilter_c1 <- summarise(group_by(dw2, Musicianship, Familiarity), n = n())
knitr::kable(Sfilter_c1)

# Indian
di2 <- di[!(di$ResponseId %in% c('R_4PhKWMzINpHFoVy', 'R_4Y9utJ6hhxLFuDy', 'R_4eWPin0vYILik8E', 'R_4qsOSprsTyIzn7e')), ]
# Checking the loss 
Sfilter_c2 <- summarise(group_by(di2, Musicianship), n = n())
knitr::kable(Sfilter_c2)
```
##Removing Participants and writing preprocessed data for analysis
```{r remove}
# Western
dw_long2 <- dw_long[!(dw_long$ResponseId %in% c('R_2Uax5iSyFsc52pZ', 'R_3kpjZD1SAq9rBt8', 'R_3lYS6mLft5j6zOW')), ]
#Saving data
write.csv(dw_long2, file = "Westerndata_outlierremoved.csv", row.names = FALSE)

# Indian
di_long2 <- di_long[!(di_long$ResponseId %in% c('R_4PhKWMzINpHFoVy', 'R_4Y9utJ6hhxLFuDy', 'R_4eWPin0vYILik8E', 'R_4qsOSprsTyIzn7e')), ]
write.csv(di_long2, file = "Indiandata_outlierremoved.csv", row.names = FALSE)
```