---
title: "Summary Statistics"
author: "T E edited M C edited NA"
date: '2024-02-20'
output: 
  html_document:
    toc: false
    number_sections: true
    theme: flatly
    highlight: espresso
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initialise

## Libraries

```{r  libraries,warning=FALSE,message=FALSE}
library(emmeans)
library(dplyr)
library(ggplot2)
library(ggforce)
library(tidyverse)
library(ggplot2)
library(stringr)
library(lme4)

```

## Read data

Reading the outlier removed data

```{r readdata}
dw2<-read.csv('Westerndata_outlierremoved.csv',skip = 0,header = TRUE)
dw2 <- subset(dw2, select = -Familiarity)
di2<-read.csv('Indiandata_outlierremoved.csv',skip = 0,header = TRUE)
```

## Combine data

```{r combine}
dw2$Familiarity <- "Non-Indian"
di2$Familiarity <- "Indian"
dcom <- rbind(dw2, di2)
dcom <- dcom %>%filter(type != "Naturll" & type != "Naturl2") #remove testing stimuli
```

# Analysis

## Accuracy Scores of IS and NI and Effect of Learning

1. Recode type into Natural, Metrical (Basic, Complex,
SP1) or Altered (SD1)

```{r,results='asis'}
dcom$type<-factor(dcom$type)
dcom$Alteration_Type <- factor(dcom$type,levels = c("Basic", "Complex", "Naturl", "sd1", "sp1"),labels = c("Metrical", "Metrical","Natural", "Altered", "Metrical"))

```

2. Plots
```{r, echo=FALSE, visualise}
# Replace 'Indian' with 'Familiar' and 'Non-Indian' with 'Unfamiliar
dcom_acc <- dcom %>%
  mutate(Familiarity = ifelse(Familiarity == "Indian", "Familiar", Familiarity)) %>%
  mutate(Familiarity = ifelse(Familiarity == "Non-Indian", "Unfamiliar", Familiarity))

# Add a new column 'Groups' by combining 'Familiairity' and 'Musicianship'
dcom_acc <- mutate(dcom_acc, Groups = paste(Familiarity, Musicianship, sep = " "))

#Accuracy Scores: Accuracy=(mean of metrical+mean of natural) - (mean of altered)
d_accuracy <- dcom_acc %>%
  group_by(ResponseId, condition, pattern, Groups, Familiarity, Musicianship) %>%
  summarise(
    Difference_Scores = (mean(similarity[type %in% c("Basic", "Naturl", "Complex", "sp1")]) -
                        mean(similarity[type == "sd1"])))

d_accuracy2 <- d_accuracy %>%
  group_by(pattern, Groups, condition) %>%
  summarise(M2 = mean(Difference_Scores), 
            n = n(), 
            SE2 = sd(Difference_Scores) / sqrt(n))
#plot difference scores
pd <- position_dodge(width = 0.8) 
d_accuracy2 <- d_accuracy %>%
  group_by(pattern, Groups, condition) %>%
  summarise(M2 = mean(Difference_Scores), 
            n = n(), 
            SE2 = sd(Difference_Scores) / sqrt(n))
g_acc1 <- ggplot(d_accuracy2, aes(x = Groups, y = M2, fill = pattern)) +
  geom_col(position = pd, width = 0.7) +
  #geom_errorbar(aes(ymin = M2 - SE2, ymax = M2 + SE2), width = 0.2, position = pd) +
  facet_wrap(~ condition, scales = "free") +
  ylab("Accuracy Scores") +  
  xlab("Participant Groups") +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(min(d_accuracy2$M2) * 1.1, max(d_accuracy2$M2) * 1.1)  # Adjust ylim

#print(g_acc1)

#Now editing the graph:
# Rename levels of the condition factor for the graph
d_accuracy3<- d_accuracy2 %>%
  mutate(condition = ifelse(condition == "Base", "A. Base", condition)) %>%
  mutate(condition = ifelse(condition == "Test", "B. Test", condition))%>%
mutate(pattern = ifelse(pattern == "Rupak", "Rupaktal (NI)", pattern))%>%
mutate(pattern = ifelse(pattern == "Teental", "Teental (ISO)", pattern))

# Plot with updated facet labels
g_acc2 <- ggplot(d_accuracy3, aes(x = Groups, y = M2, fill = pattern)) +
  geom_col(position = pd, width = 0.7) +
  facet_wrap(~ condition, scales = "free") +
  ylab("Accuracy Scores") +  
  xlab("Participant Groups") +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylim(min(d_accuracy2$M2) * 2.5, max(d_accuracy2$M2) * 1.5)+# Adjust ylim
  labs(fill = "Pattern")

print(g_acc2)

```
Statistics:

```{r, analysis accuracy}
#1. Base

acc_base <- d_accuracy %>%
  filter(condition == "Base")
lm_acc1 <- lm(Difference_Scores ~ pattern * Familiarity*Musicianship, data = acc_base)
emm_acc <- emmeans (lm_acc1,~  pattern*Familiarity*Musicianship )
pairwise_acc <- pairs(emm_acc, by = , adjust = "bonferroni")
pairwise_summary_acc <- summary(pairwise_acc)

#2. Test

acc_Test <- d_accuracy %>%
  filter(condition == "Test")
lm_acc2 <- lm(Difference_Scores ~ pattern * Familiarity*Musicianship, data = acc_Test)
emm_acc2 <- emmeans (lm_acc2,~  pattern*Familiarity*Musicianship )
pairwise_acc2 <- pairs(emm_acc2, by = , adjust = "bonferroni")
pairwise_summary_acc2<- summary(pairwise_acc2)

#overall for effect of short term learning
m2 <- lmer(Difference_Scores ~ pattern * Familiarity * Musicianship  * condition + (1 | ResponseId), data = d_accuracy)
emm_acc3 <- emmeans(m2, ~  pattern + Familiarity + condition + pattern + Musicianship : Familiarity + pattern : condition + Musicianship : condition )
jt_acc3<-joint_tests(emm_acc3)
print(knitr::kable(jt_acc3))
pairwise_acc3 <- pairs(emm_acc3, by = , adjust = "bonferroni")
pairwise_summary_acc3 <- summary(pairwise_acc3)

```
One could add the p-values to the plot above using annotate function.

## Change in Metrical Structure on Similarity perception
plots
```{r,echo=FALSE,fig.width=10,fig.height=6}
vis_meter1 <- dcom %>%
  filter(condition != "Base") %>%
  mutate(Familiarity = ifelse(Familiarity == "Indian", "Familiar",
                              ifelse(Familiarity == "Non-Indian", "Unfamiliar", Familiarity)),
         Groups = paste(Familiarity, Musicianship, sep = " "),
         Alteration_Type = factor(Alteration_Type, levels = c("Natural", "Metrical", "Altered")),
         pattern = ifelse(pattern == "Rupak", "Rupaktal (NI)", pattern),
         pattern = ifelse(pattern == "Teental", "Teental (ISO)", pattern)) %>%
  group_by(pattern, Alteration_Type) %>%
  summarise(M = mean(similarity), n = n(), SE = sd(similarity) / sqrt(n), .groups = 'drop')

g_meter1 <- ggplot(vis_meter1, aes(x = Alteration_Type, y = M, fill = Alteration_Type)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.7) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.2, position = position_dodge(width = 0.7)) +
  facet_wrap(~ pattern, scales = "fixed") +
  ylab("Similarity Rating") +
  xlab("Meter Alteration-Types") +
  #ggtitle("Similarity Ratings After Short-Term Learning") +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) +
  #labs(fill = "Alteration Types") +
  scale_fill_brewer(palette = "Set2")+
  ylim(0, 7.5)
g_meter1
#only Base to check
vis_meter1.2 <- dcom %>%
  filter(condition != "Test") %>%
  mutate(Familiarity = ifelse(Familiarity == "Indian", "Familiar",
                              ifelse(Familiarity == "Non-Indian", "Unfamiliar", Familiarity)),
         Groups = paste(Familiarity, Musicianship, sep = " "),
         Alteration_Type = factor(Alteration_Type, levels = c("Natural", "Metrical", "Altered"))) %>%
  group_by(pattern, Alteration_Type) %>%
  summarise(M = mean(similarity), n = n(), SE = sd(similarity) / sqrt(n), .groups = 'drop')
g_meter1.2 <- ggplot(vis_meter1.2, aes(x = Alteration_Type, y = M, fill = Alteration_Type)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.7) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.2, position = position_dodge(width = 0.7)) +
  facet_wrap(~ pattern, scales = "fixed") +
  ylab("Similarity Rating") +
  xlab("Meter Alteration Types") +
  ggtitle("Similarity Ratings by Alteration Type and Pattern") +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) +
  #labs(fill = "Alteration Types") +
  scale_fill_brewer(palette = "Set2")+
  ylim(0, 7.5)
g_meter1.2

#Participant vise comparison for only test condition
vis_meter2 <- dcom %>%
  filter(condition != "Base") %>%
  mutate(
    Familiarity = ifelse(Familiarity == "Indian", "Familiar",
                         ifelse(Familiarity == "Non-Indian", "Unfamiliar", Familiarity)),
    Groups = paste(Familiarity, Musicianship, sep = " "),
    Alteration_Type = factor(Alteration_Type, levels = c("Natural", "Metrical", "Altered")),
    pattern = ifelse(pattern == "Rupak", "Rupaktal (NI)", pattern),
    pattern = ifelse(pattern == "Teental", "Teental (ISO)", pattern)
  ) %>%
  group_by(Groups, pattern, Alteration_Type) %>%
  summarise(M = mean(similarity), n = n(), SE = sd(similarity) / sqrt(n), .groups = 'drop')

g_meter2 <- ggplot(vis_meter2, aes(x = Groups, y = M, fill = Alteration_Type)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.7) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.2, position = position_dodge(width = 0.7)) +
  facet_wrap(~ pattern, scales = "fixed") +
  ylab("Similarity Ratings") +
  xlab("Participant Groups") +
  #ggtitle("Similarity Ratings by Alteration Type and Pattern") +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) +
  labs(fill = "Alteration Types") +
  scale_fill_brewer(palette = "Set2") +
  ylim(0, 7.5)
g_meter2

g_meter3 <- ggplot(vis_meter2, aes(x = Alteration_Type, y = M, fill = Groups)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.7) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.2, position = position_dodge(width = 0.7)) +
  facet_wrap(~ pattern, scales = "fixed") +
  ylab("Similarity Ratings") +
  xlab("Alteration_Type") +
  ggtitle("Similarity Ratings by Alteration Type and Pattern") +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) +
  labs(fill = "Groups") +
  scale_fill_brewer(palette = "Set2") +
  ylim(0, 7.5)
g_meter3

```
Statistics:
```{r, analysis meter}

#Overall comparison of alteration types for only Test Learning Condition

meter_overall <- dcom %>%
  filter(condition != "Base")
lm_met1 <- lmer(similarity ~ pattern * Familiarity* Musicianship*Alteration_Type + (1 | ResponseId), data = meter_overall)
emm_met1 <- emmeans (lm_met1,~  + pattern +Familiarity +   Musicianship +Alteration_Type+ pattern : Alteration_Type)
jt_met1<-joint_tests(emm_met1)
print(knitr::kable(jt_met1))
#1. for overall pattern and alteration type
emm_met_overall<- emmeans (lm_met1,~  pattern*Alteration_Type )
pairwise_met1 <- pairs(emm_met_overall, by = , adjust = "bonferroni")
pairwise_summary_met1 <- summary(pairwise_met1)
#2. for all factors
pairwise_met2 <- pairs(emm_met1, by = , adjust = "bonferroni")
pairwise_summary_met2 <- summary(pairwise_met2)



```
